---
title: "data_607-week_03"
author: "Nicholas Chung"
date: "09/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# load packages
library(stringr)
library(XML)
library(RCurl)
library(tau)
```

## week 3 assignment

##### Please deliver links to an R Markdown file (in GitHub and rpubs.com) with solutions to problems 3 and 4 from chapter 8 of Automated Data Collection in R.  Problem 9 is extra credit.
```{r}
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"

# Extract information
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name
phone <- unlist(str_extract_all(raw.data, "\\(?(\\d{3})?\\)?(-| )?\\d{3}(-| )?\\d{4}"))
phone
data.frame(name = name, phone = phone)
```

##### 3.1
##### Use the tools of this chapter to rearrange the vector so that all elements conform to the standard "first_name last_name".

##### Answer:
```{r}
# use grep to find strings with commas to identify names with "last, first" convention
first_lasts <- grep(",", name, value=TRUE)
first_lasts

# split each string by comma & remove leading whitespace
name_a <- str_trim(unlist(strsplit(first_lasts[1], ",")), side = "left")
name_1 <- paste(name_a[2], name_a[1])
name_b <- str_trim(unlist(strsplit(first_lasts[2], ",")), side = "left")
name_2 <- paste(name_b[2], name_b[1])
first_last <- c(name_1, name_2)

# replace names in original vector
name[str_detect(name,"[[:alpha:]],")] <- first_last
name
```

##### 3.2. 
##### Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.).

##### Answer:
```{r}
# look for strings with a period following the second or third character
title <- str_detect(name, "[[:alpha:]]{2,3}\\.")
title

# show matching names
name[title]
```

##### 3.3. 
##### Construct a logical vector indicating whether a character has a second name.

##### Answer:
```{r}
# assumption: "second name" == "middle name"
# remove titles and trailing spaces
no_titles <- str_trim(str_remove(name, "[[:alpha:]]{2,3}\\."))
no_titles

# find names with strings with leading and trailing spaces to indicate presence of a middle name
grep("[[:space:]]+[[:alpha:]]+[[:space:]]", no_titles, value=TRUE)
```

##### 4.1
##### Describe the type of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

##### Answer:
```{r}
#1 [0-9]+\\$ 
# match any digit one or more times, followed by a '$'
grep(pattern = "[0-9]+\\$", "2756$", value = TRUE)

#2 \\b[a-z]{1,4}\\b
# match 'b', followed by any lower case ASCII letter at least 1 time, but not more than 4 times, followed by a 'b' 
grep(pattern = "\\b[a-z]{1,4}\\b", "blab", value = TRUE )

#3 .*?\\.txt$
# match an optional '.' zero or one time, then a '.txt' at the end 
grep(pattern = ".*?\\.txt$", ".foo.txt", value = TRUE )

#4 \\d{2}/\\d{2}/\\d{4} 
# match digits exactly twice, forward slash, digits exactly twice, forward slash, then digits four times
grep(pattern = "\\d{2}/\\d{2}/\\d{4}", "09/15/2019", value = TRUE )

#5 <(.+?)>.+?</\\1> 
# match any optional character immediately preceded by any character wrapped by less-than and greater-than, followed by any optional character immediately preceded by any character, followed by a forward-slash that is followed by the first parenthesis-delimited expression    
grep(pattern = "<(.+?)>.+?</\\1>", "It is possible to parse HTML markup <strong>formatted</strong> <em>text</em>.", value = TRUE )
```

##### 9.1
##### The following code hides a secret message. Crack it with R and regular expressions. Hint: Some of the characters are more revealing than others! The code snippet is also available in the materials www.r-datacollection.com.

##### Answer:
```{r}
code <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"
str_replace_all(code, "[[:lower:]]|[[:digit:]]", "")
```


