---
title: "data_607-week_03"
author: "Nicholas Chung"
date: "09/22/2019"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# load packages
library(stringr)
library(XML)
library(RCurl)
library(tau)
library(dplyr)
```

# 1. question 
##### In this project, you’re given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:
##### Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents
##### For the first player, the information would be:
##### Gary Hua, ON, 6.0, 1794, 1605
##### 1605 was calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and dividing by the total number of games played.

# 2. methodology
##### In the most general strokes, my approach to this project was comprised of three parts:
##### 2-1. data loading
##### 2-2. regex matching
##### 2-3. referencing and joining 
##### In "data loading", I tried several different approaches, including using the "scan" and "readLines" methods, but settled for "read.delim" to handle the delimiter explicitly.
##### In "regex matching", I used regex to match relevant values while preserving empty values, which were necessary for calculating number of games each player played.
##### In "referencing and joining", I first recreated the original data structure as a dataframe as an intermediate frame of reference. Then I extracted a subset to simplify mapping and validate additional column values before producing the desired output and writing to file. 

# 2-1. data loading
```{r}
# read pipe delimited text file in [table format|https://www.rdocumentation.org/packages/psych/versions/1.8.12/topics/read.file], skipping the 1st line
raw_txt_a = read.delim("tournamentinfo.txt", sep = "|", skip = 1) # read txt file
head(raw_txt_a)
```

```{r}
# read existing headers
names(raw_txt_a)

# read first two lines
head(raw_txt_a, n = 2) 

# rename headers
h_txt <- rename(raw_txt_a, pair_num = Pair, player_name = Player.Name, total_points = Total, round_1 = Round, round_2 = Round.1, round_3 = Round.2, round_4 = Round.3, round_5 = Round.4, round_6 = Round.5, round_7 = Round.6) 
head(h_txt) 
```

```{r}
# remove first row
rh_txt <- h_txt[-c(1), ]
head(rh_txt)
```

```{r}
# match rows with dashes
head(str_subset(rh_txt$pair_num, "[-]"))
```

```{r}
# remove rows with dashes
drh_txt <- subset(rh_txt, pair_num!=str_subset(rh_txt$pair_num, "[-]"))
head(drh_txt)
```

# 2-2. regex matching
```{r}
 # match player number
player_numbers <- trimws(str_subset(drh_txt$pair_num, "[[:digit:]]")) # do I need to change to int?

# check for completeness against expected count
length(player_numbers) 
```

```{r}
# match player name
names <- trimws(str_extract_all(drh_txt$player_name, "^[A-Z\\- ]+$", simplify = TRUE)) 

# remove empty strings
names <- names[names != ""]

# check for completeness against expected count
length(names)
```

```{r}
# match player state
player_state <- trimws(str_subset(drh_txt$pair_num, "[[:alpha:]]"))

# check for completeness against expected count
length(player_state)
```

```{r}
# match total points
total_points <- trimws(str_subset(drh_txt$total_points, "\\d(\\.\\d)"))

# check for completeness against expected count
length(total_points)
``` 

```{r}
# match player pre-rating
# " 15445895 / R: 1794   ->1817" 
pre_rating <- trimws(unlist(str_extract_all(drh_txt$player_name, "[:space:]\\d{3,4}", simplify = TRUE))[,2])

# remove empty strings
pre_rating <- pre_rating[pre_rating != ""]

# check for completeness against expected count
length(pre_rating)
```

# 2-3. referencing and joining
```{r}
# create matrix from relevant data
player_tbl <- cbind(player_numbers, player_state, names, pre_rating, total_points) # create matrix

# convert matrix to dataframe
player_tbl <- data.frame(player_tbl)
head(player_tbl)
```

```{r}
# replicate gameplay data
games <- subset(drh_txt, select = round_1:round_7)

# create matrix excluding irrelevant rows
games <- games[seq(1, nrow(games), 2), ] 

# extract player numbers from matrix and transpose
games <- t(apply(games, 1, function(x) str_extract_all(x, "[0-9]{1,2}", simplify = TRUE)))

# store matrix as dataframe
games_tbl <- data.frame(games)

# reorder dataframe by row number
rownames(games_tbl) <- seq(length = nrow(games_tbl))

# optionally, rename headers for ease of reference
games_tbl <- rename(games_tbl, round_1 = X1, round_2 = X2, round_3 = X3, round_4 = X4, round_5 = X5, round_6 = X6, round_7 = X7) 

# produce cleaned version of original tournamentinfo table for reference
tournament_info <- merge(x = player_tbl, y = games_tbl, by = 0)
head(tournament_info)
```
```{r}
# replace player numbers in games columns with relevant pre-rating scores

# create key, value list pair
key <- c(as.character(tournament_info$player_numbers))
value <- c(as.character(tournament_info$pre_rating))
# create map from keys and values 
map <- setNames(value, key)

# map new values to dataframe
games_tbl[] <- map[unlist(games_tbl)]

# convert new values to integer class
games_tbl[, ] <- lapply(games_tbl[, ], as.integer)

# sum total score per row while accounting for missing values
games_tbl$total_score <- rowSums(games_tbl[1:7], na.rm = TRUE)

# count number of games played by discounting missing values
games_tbl$n_games <- rowSums(!is.na(games_tbl[1:7])) 

# store calculation of average rating per row in the new column
games_tbl[, "opp_avg_score"] <- round(games_tbl[, "total_score"] / games_tbl[, "n_games"])
head(games_tbl)
```

# 3. solution
```{r}
# create desired output

# store outputs in desired classes as vectors
player_num <- as.vector(tournament_info$player_numbers)
player_nom <- as.vector(tournament_info$names)
player_state <- as.vector(tournament_info$player_state)
total_points <- as.vector(tournament_info$total_points)
pre_rating <- as.integer(as.vector(tournament_info$pre_rating))
opponent_rating <- as.vector(games_tbl$opp_avg_score)

# create desired dataframe from output vectors
output_tbl <- data.frame(player_num, player_nom, player_state, total_points, pre_rating, opponent_rating)
output_tbl

# write dataframe to CSV
write.table(output_tbl, file = "tournamentinfo.csv", row.names=FALSE,sep = ",")
```

