---
title: "cuny_msds-math_bridge_hw-1"
author: "Nicholas Chung"
date: "7/20/2019"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Homework 1 

### Derivatives
Find the derivatives with respect to x of the following:
  
#### 1Q
\[
  F(x|x \geq 0) = 1 - e^{\lambda*x}
\]    

#### 1A
```{r}
library(Deriv)

# store function
f1=function(x)1-(exp(-lambda*x))

# derivative of function
ans=Deriv(f1)
ans
```

#### 2Q
\[
  F(x|b>a) = \frac{x - a}{b-a}
\]    

#### 2A
```{r}
# store function
f2=function(x)(x-a)/(b-a)

# derivative of function
ans=Deriv(f2)
ans
```

#### 3Q
\[
  F(x|a<x \leq c \leq b) = \frac{(x-a)^2}{(b-a)(c-a)}
\]    

#### 3A
```{r}
# store function
f3=function(x)(x-a)^2/((b-a)(c-a))

# derivative of function
ans=Deriv(f3)
ans
```

#### 4Q
\[
  F(x|a<x \leq c < x < b) = 1 - \frac{(b-x)^2}{(b-a)(c-a)}
\]    

#### 4A
```{r}
# store function
f4=function(x)1-((b-x)^2/((b-a)(c-a)))

# derivative of function
ans=Deriv(f4)
ans
```

### Integrals
Solve the following definite and indefinite integrals

#### 5Q
\[
  \int_{0}^{10}3x^{3}dx
\]    

#### 5A
```{r}
# store function
f5=function(x)(3*(x^3))

# integrate definite integral
ans=integrate(Vectorize(f5),0,10)
ans
```

#### 6Q
\[
  \int_{0}^{x}x\lambda e^{-\lambda*x} dx
\]    

#### 6A
```{r}
library(mosaicCalc)

# integrate indefinite integral
ans=antiD(x*lambda*exp(-lambda*x)~x)
ans
```

#### 7Q
\[
  \int_{0}^{.5}\frac{1}{b-a}dx
\]    

#### 7A
```{r}
# assign values to variables a & b
a<-0
b<-1

# store function
f7=function(x)(1/(b-a))

# integrate definite integral
ans=integrate(Vectorize(f7),0,.5)
ans
```

#### 8Q
\[
  \int_{0}^{x}x\frac{1}{\Gamma(\alpha)\beta^\alpha}x^{\alpha-1}e^{-\beta x}dx
\]    

#### 8A
```{r}
# integrate indefinite integral
ans=antiD(x^{alpha+1}/gamma*alpha*beta^{alpha}*e^{beta}~x)
ans
```

### Linear Algebra
With the following matrix,

\[
\quad
x=\begin{bmatrix}
1 & 2 & 3 \\
3 & 3 & 1 \\
4 & 6 & 8
\end{bmatrix}
\]

#### 9Q
Invert it using Gaussian row reduction.    

#### 9A
```{r}
library(matlib)
# store matrix
A9<-rbind(c(1,2,3),c(3,3,1),c(4,6,8))
A9

# determine whether matrix is solvable
gaussianElimination(A9, numeric(3))

# find inverse matrix by elimination
gaussianElimination(A9, diag(3))
ans = inv(A9)
ans

# validate above solution by finding inverse using solve method
B9=solve(A9)
B9
```

#### 10Q
Find the determinant.    

#### 10A
```{r}
# find determinant of matrix
ans=det(A9)
ans
```

#### 11Q
Conduct LU decomposition.    

#### 11A
```{r}
library(pracma)

# use lu() method for LU decomposition
ans=lu(A9)
ans
```

#### 12Q
Multiply the matrix by its inverse.    

#### 12A
```{r}
# multiply matrix A9 and its inverse B9
ans=A9%*%B9
ans
```